Generalized Tractability

### 2023-04-11 Older_versions/GeneralizedTractability_230310.tex
Updates from Overleaf

\begin{definition}[$T$-Quasi-polynomial tractability]
We say that the problem is \emph{$T$- quasi-polynomial tractable} if there exists positive $C_{\vp}$ such that
   %\[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(0,d,\vp))(T(\varepsilon^{-1},1,\vp)\}\]
   \[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(\varepsilon^{-1},d,\vp)\}\]
\end{definition}

\begin{theorem}\label{thm:qpt}
A problem is $T$-quasi-polynomial tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ such that
\begin{equation}\label{thmqpt}
M_{\vp} := \sup_{d\in\mathbb{N}}\sum_{i=1}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d,\vp)}
<\infty
\end{equation}
%where $\Bar{p} = -p(1+\log(T(0,d,\vp))$.
%If it holds then 
%\[n(\varepsilon,d)\leq 1+T(0,d^\tau,\vp)+ T(0,d^\tau,\vp)T(\varepsilon^{-1},1,\vp)\]
\end{theorem}
\begin{proof}
\textbf{Necessity Condition:}\\
Suppose that we have $T$-quasi-polynomial tractability, i.e., for some $\vp\geq 0$, there exists positive $C_{\vp}$ such that
%\[\comp(\varepsilon,d)\le C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}\quad\quad \forall\varepsilon >0, d\in \mathbb{N}\] 
\[
\comp(\varepsilon,d)\le C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}
\quad\quad \forall\varepsilon >0, d\in \mathbb{N}
\] 
Since $\lambda_{i,d}$ are non-increasing, we have
\[
\lambda_{\left\lfloor C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}\right\rfloor+1,d}\leq \varepsilon^2
\]
Given that the estimate holds for all $\varepsilon> 0$, we assume that $\varepsilon \in (0, 1]$.\\

If we vary $\varepsilon\in (0,1]$, we have that
\[
j(\varepsilon,d,\vp) = \left\lfloor C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}\right\rfloor+1
\] which attains the values $j = j_d,j_{d+1},\dots$, where
\[
j(\infty,d,\vp)=j_d = \lfloor C_{\vp}\exp\{ T(0,d,\vp)\}\rfloor + 1
\]
Furthermore, we have
\[j \leq C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}+1 \leq C_{\vp}\exp\{T(\lambda^{-1}_{j,d},d,\vp)\}+1 \]
This implies that for $\tau>1$, we have
%\[\left(\log\frac{j-1}{C_{\vp}}\right)^\tau \leq \left(T(\varepsilon^{-1},d,\vp)\right)^\tau \leq K_{\vp,\tau}T(\lambda_{j,d},d,\tau\vp) \]
we have 
%\[\frac{1}{T(\lambda_{i,d}^{-1},d,\vp)} \leq \frac{K_{\vp,\tau}}{\left(\log\frac{j-1}{C_{\vp}}\right)^\tau}\]
\end{proof}

@@

\noindent \textbf{Necessary Condition:}\\
Suppose that the problem is  sub-$h \circ T$ tractable. Then, for any $c>0$ there exists a positive integer $V_c$ such that  
\[
	\comp(\varepsilon,d)\le 
	\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor \qquad \forall \varepsilon^{-1} + d \ge V_c.
\]
By the definition of complexity in \eqref{eq:comp_def}, it follows that 
\[
	\lambda_{n(\varepsilon,d,c),d} \le \varepsilon \quad \forall \varepsilon^{-1} + d \ge V_c, \qquad \text{where }
	n(\varepsilon,d,c):=\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor.
\]
Since $h(cT(\cdot,d))$ is non-decreasing, it follows that
\begin{equation*}
		n(\varepsilon,d,c)\le  h(cT(\varepsilon^{-1},d)) \le  h(cT(\lambda^{-1}_{n(\varepsilon,d,c),d},d)) \qquad  \quad \forall \varepsilon^{-1} + d \ge V_c.
\end{equation*}

For a fixed $d$, define $\varepsilon^{-1}_{\max,d} = \max(V_c - d,\lambda_{\max}^{-1})$.  the largest the largest $\varepsilon$ can be and still satisfy $ \varepsilon^{-1} + d \ge V_c$ is $\max(V_c - d,1)$.   Let 
\begin{equation} \label{eq:nstardef}
	\varepsilon^{-1}_{\max,d} := \max(V_c - d,\lambda_{\max}^{-1}), \qquad n^*(d,c) : = n(\varepsilon^{-1}_{\max,d},d,c) \le h(cT(\varepsilon^{-1}_{\max,d},d)).
\end{equation} 
By varying $\varepsilon^{-1}$ in the interval $[\varepsilon^{-1}_{\max,d},\infty)$,  the integer $n(\varepsilon,d,c)$ attains all values greater than or equal to $n^*(d,c)$.  Thus, 
\begin{align}
		n &\le h(cT(\lambda^{-1}_{n,d},d)) \qquad  \forall n \ge n^*(d,c), \label{eq:subhTneq}\\
		\nonumber
		\frac{1}{h(2 cT(\lambda^{-1}_{n,d},d)) } & \le \frac{1}{[h(cT(\lambda^{-1}_{n,d},d))]^{2}}  \le \frac 1{n^2}   \qquad  \forall n \ge n^*(d,c) \qquad \text{by \eqref{eq:h_cond}}. \\
		\intertext{Combining this inequality with the upper bound on $n^*(d,c)$ yields}
		\nonumber
		S_{2c} & =\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} \\
		\nonumber
		&  = \sup_{d\in\naturals} \left [
		\sum_{i=1}^{n^*(d,c)-1} \frac1{h(2cT(\lambda_{i,d}^{-1},d))}
		+ \sum_{i=n^*(d,c)}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} 
		\right]\\
		 \nonumber
		 & \le \sup_{d\in\naturals} \left [
\frac{n^*(d,c)-1}{h(2cT(\lambda_{\max}^{-1},d))}
+ \sum_{i=1}^\infty \frac1{n^2}
\right] 
 \qquad \text{by \eqref{eq:subhTneq}}\\ 
 \label{eq:lastone}
		 & \le \sup_{d\in\naturals} \left [
		 \frac{ h(cT(\varepsilon^{-1}_{\max,d},d))}{h(2cT(\lambda_{\max}^{-1},d))}
		\right] + \frac{\pi^2}{6}
		\qquad \text{by \eqref{eq:nstardef}.}
\end{align}

Note that 
\[
\varepsilon^{-1}_{\max,d} = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^* : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in it's arguments, this implies that 
\begin{align*}
    \sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max,d},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max,d},d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max,d},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*+1, \ldots \}} \,
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & < \infty \qquad \text{since }h \ge 1.
\end{align*}
Combining this bound with \eqref{eq:lastone} establishes the finiteness of $S_2c$ and completes the proof of the necessary condition.

>>>>

\begin{definition}[$T$-Quasi-polynomial tractability]
We say that the problem is \emph{$T$- quasi-polynomial tractable} if there exists positive $C_{\vp}$ such that
   \[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(0,d,\vp))(T(\varepsilon^{-1},1,\vp)\}\]
  % \[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(\varepsilon^{-1},d,\vp)\}\]
\end{definition}

\begin{theorem}\label{thm:qpt}
A problem is $T$-quasi-polynomial tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ such that
\begin{equation}\label{thmqpt}
M_{\vp} := \sup_{d\in\mathbb{N}}\sum_{i=1}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d,\vp)}
<\infty
\end{equation}
%where $\Bar{p} = -p(1+\log(T(0,d,\vp))$.
%If it holds then 
%\[n(\varepsilon,d)\leq 1+T(0,d^\tau,\vp)+ T(0,d^\tau,\vp)T(\varepsilon^{-1},1,\vp)\]
\end{theorem}
\begin{proof}
\textbf{Necessity Condition:}\\
Suppose that we have $T$-quasi-polynomial tractability, i.e., for some $\vp\geq 0$, there exists positive $C_{\vp}$ such that
\[\comp(\varepsilon,d)\le C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}\quad\quad \forall\varepsilon >0, d\in \mathbb{N}\] 
%\[\comp(\varepsilon,d)\le C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}\quad\quad \forall\varepsilon >0, d\in \mathbb{N}\] 
Since $\lambda_{i,d}$ are non-increasing, we have
\[
\lambda_{\left\lfloor C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}\right\rfloor+1,d}\leq \varepsilon^2
\]
Given that the estimate holds for all $\varepsilon> 0$, we assume that $\varepsilon \in (0, 1]$.\\

If we vary $\varepsilon\in (0,1]$, we have that
\[
j(\varepsilon,d,\vp) = \left\lfloor C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}\right\rfloor+1
\] which attains the values $j = j_d,j_{d+1},\dots$, where
\[
j(\infty,d,\vp)=j_d = \lfloor C_{\vp}\exp\{ T(0,d,\vp)\}\rfloor + 1
\]
Furthermore, we have
\[j \leq C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}+1 \leq C_{\vp}\exp\{(T(0,d,\vp)(T(\lambda^{-1}_{j,d},1,\vp))\}+1 \quad \forall j \geq j_d \]
This implies that for $\tau>1$, we have
\[
\left(\frac{j-1}{C_{\vp}}\right)^{\tau/T(0,d,\vp)} \leq \exp\{T(\lambda^{-1}_{j,d},1,\vp)\}
\]
\[
\exp\{-\tau T(\lambda^{-1}_{j,d},1,\vp\} \leq \left(\frac{C_{\vp}}{j-1}\right)^{\tau/T(0,d,\vp)}\Rightarrow
\exp\{-\tau T(\lambda^{-1}_{j,d},1,\vp)(T(0,d,\vp)\} \leq C_{\vp}^\tau (j-1)^{-\tau}
\]

%\[\frac{1}{T(\lambda_{i,d}^{-1},d,\vp)} \leq \frac{K_{\vp,\tau}}{\left(\log\frac{j-1}{C_{\vp}}\right)^\tau}\]

\textbf{Sufficient condition:} Suppose that (\ref{}) holds for some $\vp \geq 0$. We need to show that for some $C_{\vp}>0$ we have 
\[
\text{comp}(\varepsilon,d) \leq C_{\vp}\exp\{T(0,d,\vp)T(\varepsilon^{-1},1,\vp)\} \quad \forall \varepsilon > , d \in \mathbb{N}
\] Let
\end{proof}

@@

\noindent \textbf{Necessary Condition:}\\
Suppose that the problem is  sub-$h \circ T$ tractable. Then, for any $c>0$ there exists a positive integer $V_c$ such that  
\[
	\comp(\varepsilon,d)\le 
	\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor \qquad \forall \varepsilon^{-1} + d \ge V_c.
\]
By the definition of complexity in \eqref{eq:comp_def}, it follows that 
\[
	\lambda_{n(\varepsilon,d,c),d} \le \varepsilon \quad \forall \varepsilon^{-1} + d \ge V_c, \qquad \text{where }
	n(\varepsilon,d,c):=\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor.
\]
Since $h(cT(\cdot,d))$ is non-decreasing, it follows that
\begin{equation}
		n(\varepsilon,d,c)\le  h(cT(\varepsilon^{-1},d)) \le  h(cT(\lambda^{-1}_{n(\varepsilon,d,c),d},d)) \qquad  \quad \forall \varepsilon^{-1} + d \ge V_c. \label{eq:nedcbound}
\end{equation}

For a fixed $c$ and $d$, define
\begin{equation} \label{eq:nstardef}
	\varepsilon^{-1}_{\max}(d,c) := \max(V_c - d,\lambda_{\max}^{-1}), \qquad n^*(d,c) : = n(\varepsilon^{-1}_{\max}(d,c),d,c) \le h(cT(\varepsilon^{-1}_{\max}(d,c),d)).
\end{equation} 
By varying $\varepsilon^{-1}$ in the interval $[\varepsilon^{-1}_{\max}(d,c),\infty)$,  the integer $n(\varepsilon,d,c)$ attains all values greater than or equal to $n^*(d,c)$.  Thus, by \eqref{eq:nedcbound},
\begin{align}
		n &\le h(cT(\lambda^{-1}_{n,d},d)) \qquad  \forall n \ge n^*(d,c), \label{eq:subhTneq}\\
		\nonumber
		\frac{1}{h(2 cT(\lambda^{-1}_{n,d},d)) } & \le \frac{1}{[h(cT(\lambda^{-1}_{n,d},d))]^{2}}  \le \frac 1{n^2}   \qquad  \forall n \ge n^*(d,c) \qquad \text{by \eqref{eq:h_cond}}. \\
		\intertext{Combining this inequality with the upper bound on $n^*(d,c)$ yields}
		\nonumber
		S_{2c} & =\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} \\
		\nonumber
		&  = \sup_{d\in\naturals} \left [
		\sum_{i=1}^{n^*(d,c)-1} \frac1{h(2cT(\lambda_{i,d}^{-1},d))}
		+ \sum_{i=n^*(d,c)}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} 
		\right]\\
		 \nonumber
		 & \le \sup_{d\in\naturals} \left [
\frac{n^*(d,c)-1}{h(2cT(\lambda_{\max}^{-1},d))}
+ \sum_{i=1}^\infty \frac1{n^2}
\right] 
 \qquad \text{by \eqref{eq:subhTneq}}\\ 
 \label{eq:lastone}
		 & \le \sup_{d\in\naturals} \left [
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
		\right] + \frac{\pi^2}{6}
		\qquad \text{by \eqref{eq:nstardef}.}
\end{align}

Note that 
\[
\varepsilon^{-1}_{\max}(d,c) = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^*_c : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*_c+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in it's arguments, this implies that 
\begin{align*}
    \MoveEqLeft{\sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}}\\
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d^*_c+1))}
  \right \} \\
  & < \infty.
\end{align*}
Combining this bound with \eqref{eq:lastone} establishes the finiteness of $S_{2c}$ and completes the proof of the necessary condition.
